Predicting information integration in pragmatic word learning across development

### 2020-09-23 papers/full_paper/Supplementary_information.Rmd
revision sent

## Model parameters

@@

For children, we used the data from Experiment 6 to model the intercepts ($\beta^\rho_{0,j}$) and slopes ($\beta^\rho_{1,j}$) that best described the developmental trajectory in the data for each of the two ($j$) conditions. As for the speaker optimality parameter, this allowed us to generate object prior distributions in the cognitive model that were sensitive to the child's age. We used a simple Bayesian logistic regression to find the intercept and slope (MAP of posterior distribution) that best described children's performance in the two conditions of experiment 6. The prior distribution was $\beta^\rho_0 ,\beta^\rho_1 \sim \text{Uniform}(-2,2)$ for the intercept and slope. We collected samples from four independent MCMC chains, collecting 6000 samples from each chain and removing the first 2000 for burn-in. In Experiment 7, the prior probability for an object was computed by taking the intercept for the respective condition ($j$), adding the slope times the child's age ($i$) and then using a logistic transformation to convert the outcome into proportions: $\rho_{i,j} = \text{logistic}(\beta^\rho_{0,j} + i \cdot \beta^\rho_{1,j})$. Because these proportions corresponded to a two object scenario, they were then converted to the three object scenario via normalization in the same way as for adults. The overall distribution then depended on the alignment of information sources in the same way as it did for adults. Model code for inferring the intercept and the slope for the child study can be found in the associated online repository in the file `models.Rmd` starting line 1491.

>>>>

# Model parameters

@@

For children, we used the data from Experiment 6 to model the intercepts ($\beta^\rho_{0,j}$) and slopes ($\beta^\rho_{1,j}$) that best described the developmental trajectory in the data for each of the two ($j$) conditions. As for the speaker optimality parameter, this allowed us to generate object prior distributions in the cognitive model that were sensitive to the child's age. We used a simple Bayesian logistic regression to find the intercept and slope (MAP of posterior distribution) that best described children's performance in the two conditions of experiment 6. The prior distribution was $\beta^\rho_0 ,\beta^\rho_1 \sim \text{Uniform}(-2,2)$ for the intercept and slope. We collected samples from four independent MCMC chains, collecting 6000 samples from each chain and removing the first 2000 for burn-in. In Experiment 7, the prior probability for an object was computed by taking the intercept for the respective condition ($j$), adding the slope times the child's age ($i$) and then using a logistic transformation to convert the outcome into proportions: $\rho_{i,j} = \text{logistic}(\beta^\rho_{0,j} + i \cdot \beta^\rho_{1,j})$. Because these proportions corresponded to a two object scenario, they were then converted to the three object scenario via normalization in the same way as for adults. The overall distribution then depended on the alignment of information sources in the same way as it did for adults. Model code for inferring the parameters can be found in the associated online repository in the file `models.Rmd` starting in line 213 (adults) 1464 (children).

### 2020-09-23 papers/full_paper/manuscript.Rmd
revision sent

What someone means by an utterance is oftentimes not reducible to the words they used. It takes  pragmatic inference – context-sensitive reasoning about the speaker’s intentions - to recover the intended meaning [@grice1991studies; @levinson2000presumptive; @sperber2001relevance]. Contextual information comes in many forms. On the one hand, there is information provided by the utterance^[
We use the terms utterance, utterance-level information or utterance-level cues to capture all cues that the speaker provides for their intended meaning. This includes direct referential information in the form of pointing or gazing, semantic information in the form of conventional word meanings as well as pragmatic inferences that are licenced by the particular choice of words or actions.] itself. Competent language users expect each other to communicate in a cooperative way such that speakers produce utterances that are relevant and informative. Thus, semantic ambiguity can be resolved by reasoning about why the speaker produced this particular utterance [@grice1991studies; @clark1996using; @sperber2001relevance; @tomasello2008origins]. On the other hand, there is information provided by common ground (the body of mutually shared knowledge and beliefs between interlocutors)[@bohn2018common; @clark1996using; @clark2015common]. Because utterances are embedded in common ground, pragmatic reasoning in context always requires information integration. But how does integration proceed? And how does it develop? Verbal theories assume that information is integrated and that this process develops but do not specify how. We bridge this gap by formalizing information integration and development in a probabilistic model of pragmatic reasoning.

@@

Evidence for the use of common ground^[Common ground has traditionally been defined in recursive terms: in order to be part of common ground, some piece of information has to be not just known to both interlocutors but also known to both to be shared between them [@clark1996using]. Numerous studies probed the role of sharedness of information and found that it plays a critical role in communicative interactions [@brown2009partner; @hanna2003effects; @heller2016perspective; @mozuraitis2015privileged]. Based on this literature, one might argue that the term common ground should be restricted to describe situations in which the sharedness aspect is directly tested. However, most of this work is focused on online perspective taking. In this paper, we use the term common ground to refer to shared information that is built up over the course of an interaction - something that is supposedly easier for children [@matthews2006effect]. We assume that the consequence of a direct interaction (with matching perspectives) between the speaker and the listener is that information is not just known to both interlocutors but also assumed to be shared between them [@bohn2018common].Our main focus is on how information that is part of common ground integrates with other pragmatic information sources.] information by young children is even stronger: Common ground information guides how infants produce non-verbal gestures and interpret ambiguous utterances [@bohn2018social; @saylor2011s]. For slightly older children, common ground – in the form of knowledge about discourse novelty, preferences, and even discourse expectations – also facilitates word learning [@akhtar1996role; @bohn_le_peloquin_koymen_frank_2020; @saylor2009preschoolers; @sullivan2019discourse].

These examples, however, highlight children’s use of a single pragmatic information source or cue. Harnessing multiple – potentially competing – pragmatic cues poses a separate challenge. One aspect of this integration problem is how to balance common ground information that is built up over the course of an interaction against information gleaned from the current utterance. Much less is known about whether and how children combine these types of information. Developmental studies that look at the integration of multiple information sources more generally find that children are sensitive to multiple sources from early on [@graham2017words; @grosse201021; @khu2020preschoolers; @matthews2006effect; @nilsen2009preschoolers; @ganea2007infants]. For example, in a classic study, Nadig and Sedivy [-@nadig2002evidence] found that children rapidly integrate information provided in an utterance (a particular referring expression) with the speaker’s perspective (the objects the speaker can see). However, the information sources to be integrated in these studies are not all pragmatic in nature. Children’s ability to pick out a referent following a noun reflects their linguistic knowledge and not necessarily their ability to reason about the speaker’s intention in context. As a consequence, this work does not speak to the question of how and if listeners integrate different forms of *pragmatic* information. Thus, while many theories of pragmatic reasoning presuppose that pragmatic information sources are integrated, the nature of their relationship has typically not been specified.

@@

Cue integration is an integral part of language comprehension and learning. As such, it has been extensively studied in recent decades. The focus of this work has usually been on how adults combine perceptual, semantic or syntactic information [e.g. @tanenhaus1995integration; @kamide2003integration; @hagoort2004integration; @ozyurek2007line]. We extend the study of cue integration to pragmatics. Most importantly, however, we present a substantive theory of the integration process itself. Real world language comprehension and learning happens in socially dynamic and complex situations which inevitably require integrating multiple pragmatic information sources. The integration model provides a formal description of the (hypothetical) psychological representations that may underlie information integration. As such, our work complements theorizing about information integration in other domains of language [e.g. @fourtassi2020optimal; @mcclelland2006there; @smith2017multimodal].

@@

We did not model the social-cognitive processes that specify the probability of an object being the referent given common ground - we simply measured it empirically. As a consequence, our approach treats common ground as equivalent to more basic manipulations of contextual salience [e.g. in @frank2012predicting]. Thus, our model would not differentiate between a situation in which an object would be salient because it has been the focus of an interaction and one in which it would be more salient because it was big or colorful. Based on a process model of common ground, one could further specify how common ground information (i.e. social context) interacts with other contextual information [@degen2015wonky; @tessler2017warm]. Relatedly, our model does not take into account the important distinction for psycholinguistics, namely the difference between privileged ground vs. common ground. This distinction has been addressed computationally by Heller and colleagues [@heller2016perspective; @mozuraitis2018modeling]. In their work, they focus on how listeners identify the referent of ambiguous referring expressions. Their probabilistic model simultaneously considers the (differing) perspectives of both interlocutors and trades off between them. In principle, the model of Heller and colleagues [-@heller2016perspective] and the *integration model* could be combined with one another to address how privileged vs common ground trades off with other pragmatic information.

@@

>>>>

What someone means by an utterance is oftentimes not reducible to the words they used. It takes  pragmatic inference – context-sensitive reasoning about the speaker’s intentions - to recover the intended meaning [@grice1991studies; @levinson2000presumptive; @sperber2001relevance]. Contextual information comes in many forms. On the one hand, there is information provided by the utterance^[
We use the terms utterance, utterance-level information or utterance-level cues to capture all cues that the speaker provides for their intended meaning. This includes direct referential information in the form of pointing or gazing, semantic information in the form of conventional word meanings as well as pragmatic inferences that are licenced by the particular choice of words or actions.] itself. Competent language users expect each other to communicate in a cooperative way such that speakers produce utterances that are relevant and informative. Thus, semantic ambiguity can be resolved by reasoning about why the speaker produced this particular utterance [@grice1991studies; @clark1996using; @sperber2001relevance; @tomasello2008origins]. On the other hand, there is information provided by common ground [the body of mutually shared knowledge and beliefs between interlocutors; @bohn2018common; @clark1996using; @clark2015common]. Because utterances are embedded in common ground, pragmatic reasoning in context always requires information integration. But how does integration proceed? And how does it develop? Verbal theories assume that information is integrated and that this process develops but do not specify how. We bridge this gap by formalizing information integration and development in a probabilistic model of pragmatic reasoning.

@@

Common ground has traditionally been defined in recursive terms: in order to be part of common ground, some piece of information has to be not just known to both interlocutors but also known to both to be shared between them [@clark1996using]. Numerous studies probed the role of sharedness of information and found that it plays a critical role in communicative interactions [@brown2009partner; @hanna2003effects; @heller2016perspective; @mozuraitis2015privileged]. Based on this literature, one might argue that the term common ground should be restricted to describe situations in which the sharedness aspect is directly tested. However, most of this work is focused on online perspective taking. In this paper, we use the term common ground to refer to shared information that is built up over the course of an interaction - something that is likely easier for children [@matthews2006effect]. We assume that the consequence of a direct interaction (with matching perspectives) between the speaker and the listener is that information is mutually manifest; that is, not just known to both interlocutors but also assumed to be shared between them [@bohn2018common] and hence part of common ground. Thus, since this information is unproblematically in common ground, we can focus on how this information integrates with other pragmatic information sources.

Evidence for the use of common ground information by young children is even stronger: Common ground information guides how infants produce non-verbal gestures and interpret ambiguous utterances [@bohn2018social; @saylor2011s]. For slightly older children, common ground – in the form of knowledge about discourse novelty, preferences, and even discourse expectations – also facilitates word learning [@akhtar1996role; @bohn_le_peloquin_koymen_frank_2020; @saylor2009preschoolers; @sullivan2019discourse].

The examples discussed above, however, highlight children’s use of a single pragmatic information source or cue. Harnessing multiple – potentially competing – pragmatic cues poses a separate challenge. One aspect of this integration problem is how to balance common ground information that is built up over the course of an interaction against information gleaned from the current utterance. Much less is known about whether and how children combine these types of information. Developmental studies that look at the integration of multiple information sources more generally find that children are sensitive to multiple sources from early on [@graham2017words; @grosse201021; @khu2020preschoolers; @matthews2006effect; @nilsen2009preschoolers; @ganea2007infants]. For example, in a classic study, Nadig and Sedivy [-@nadig2002evidence] found that children rapidly integrate information provided in an utterance (a particular referring expression) with the speaker’s perspective (the objects the speaker can see). However, the information sources to be integrated in these studies are not all pragmatic in nature. Children’s ability to pick out a referent following a noun reflects their linguistic knowledge and not necessarily their ability to reason about the speaker’s intention in context. As a consequence, this work does not speak to the question of how and if listeners integrate different forms of *pragmatic* information. Thus, while many theories of pragmatic reasoning presuppose that pragmatic information sources are integrated, the nature of their relationship has typically not been specified.

@@

Cue integration is an integral part of language comprehension and learning [as well as perception more generally; @trommershauser2011sensory]. As such, it has been extensively studied in recent decades. The focus of this work has usually been on how adults combine perceptual, semantic or syntactic information [e.g. @tanenhaus1995integration; @kamide2003integration; @hagoort2004integration; @ozyurek2007line]. We extend the study of linguistic cue integration to pragmatics. Most importantly, however, we present a substantive theory of the integration process itself. Real world language comprehension and learning happens in socially dynamic and complex situations which inevitably require integrating multiple pragmatic information sources. The integration model provides a formal description of the (hypothetical) psychological representations that may underlie information integration. As such, our work complements theorizing about information integration in other domains of language comprehension [e.g. @fourtassi2020optimal; @mcclelland2006there; @smith2017multimodal].

@@

@@

We did not model the social-cognitive processes that specify the probability of an object being the referent given common ground - we simply measured it empirically. As a consequence, our approach treats common ground as equivalent to more basic manipulations of contextual salience [e.g. in @frank2012predicting]. Thus, our model would not differentiate between a situation in which an object would be salient because it has been the focus of an interaction and one in which it would be more salient because it was big or colorful. Based on a process model of common ground, one could further specify how common ground information (i.e. social context) interacts with other contextual information [@degen2015wonky; @tessler2017warm]. A further limitation of our work here is that we did not specify how common ground is integrated into the process of compositional interpretation at work in more complex sentences. This is an open challenge for future research on pragmatic inference. One possible approach would be to make inferences about relevant common ground at the level of individual lexical items and to propagate this uncertainty through compositions into larger sentences [as in e.g., @potts2016embedded].

Our model also does not take into account the important distinction for psycholinguistics, namely the difference between privileged ground vs. common ground. This distinction has been addressed computationally by Heller and colleagues [@heller2016perspective; @mozuraitis2018modeling]. In their work, they focus on how listeners identify the referent of ambiguous referring expressions. Their probabilistic model simultaneously considers the (differing) perspectives of both interlocutors and trades off between them. In principle, the model of Heller and colleagues [-@heller2016perspective] and the *integration model* could be combined with one another to address how privileged vs common ground trades off with other pragmatic information.

# Conclusion

### 2020-09-23 papers/full_paper/models.Rmd
revision sent

---
title             : "Predicting information integration in pragmatic word learning across development"
shorttitle        : "Supplementary Information"

author: 
  - name          : "Manuel Bohn"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Department of Psychology, 450 Serra Mall, Stanford, CA 94305"
    email         : "bohn@stanford.edu"
  - name          : "Michael Henry Tessler"
    affiliation   : "3"
  - name          : "Megan Merrick"
    affiliation   : "1"
  - name          : "Michael C. Frank"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Stanford University"
  - id            : "2"
    institution   : "Leipzig Research Center for Early Child Development, Leipzig University"
  - id            : "3"
    institution   : "Department of Brain and Cognitive Sciences, Massachusetts Institute of Technology"

bibliography      : ["library.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
mask              : no

class             : "man"
output            : papaja::apa6_pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, include = T)

@@

@@

#### Inferring speaker optimality parameter and setting priors

@@

### Model predictions

#### Pragmatic model

@@

```{r predictions pragmatic model ex 3 noise, include = FALSE}
# ex3_pragm_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.rsa.ex3.noise, sep='\n'),
#   data = list(allData = adult_ex3_data, allParams = model_params_ex3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex3_pragm_model_noise.rds")

@@

#### Prior only model

@@

```{r model prior only noise ex3, include = FALSE}
predictions.prior.only.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

```{r predictions prior only model ex 3 noise, include = FALSE}
# ex3_prior_only_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.prior.only.ex3.noise, sep='\n'),
#   data = list(allData = adult_ex3_data, allParams = model_params_ex3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex3_prior_only_model_noise.rds")

@@

#### Flat prior model

@@

```{r predictions flat prior model ex 3 noise, include = FALSE}
# ex3_flat_prior_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.flat.prior.ex3.noise, sep='\n'),
#   data = list(allData = adult_ex3_data, allParams = model_params_ex3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex3_flat_prior_model_noise.rds")

@@

#### Compare model predictions to data

@@

### Model comparison

@@

##### Pragmatic model

@@

```{r loglike rsa noise ex3, include = FALSE}
loglike.rsa.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

ex3_pragm_model_noise_loglike <- readRDS("../../stats/saves/ex3_pragm_model_noise_loglike.rds")%>%
  mutate(model = "pragmatic_noise")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))
```
##### Prior only model

@@

```{r loglike prior only noise ex3, include = FALSE}
loglike.prior.only.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

##### Flat prior model

@@

```{r loglike flat prior noise ex3, include = FALSE}
loglike.flat.prior.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

### Model predictions (including noise parameter)

#### Pragmatic model

```{r model rsa noise ex4, include = FALSE}
predictions.rsa.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

```{r predictions pragmatic model ex 4 noise, include = FALSE}
# ex4_pragm_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.rsa.ex4.noise, sep='\n'),
#   data = list(allData = adult_ex4_data, allParams = model_params_ex4),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex4_pragm_model_noise.rds")

@@

#### Prior only model

```{r model prior only noise ex4, include = FALSE}
predictions.prior.only.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

```{r predictions prior only model ex 4 noise, include = FALSE}
# ex4_prior_only_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.prior.only.ex4.noise, sep='\n'),
#   data = list(allData = adult_ex4_data, allParams = model_params_ex4),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex4_prior_only_model_noise.rds")

@@

#### Flat prior model

```{r model flat prior noise ex4, include = FALSE}
predictions.flat.prior.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

```{r predictions flat prior model ex 4 noise, include = FALSE}
# ex4_flat_prior_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.flat.prior.ex4.noise, sep='\n'),
#   data = list(allData = adult_ex4_data, allParams = model_params_ex4),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex4_flat_prior_model_noise.rds")

@@

### Compare model predictions to data

@@

### Model comparison

@@

##### Pragmatic model 
```{r loglike rsa noise ex4, include = FALSE}
loglike.rsa.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

##### Prior only model

```{r loglike prior only noise ex4, include = FALSE}
loglike.prior.only.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

##### Flat prior model

```{r loglike flat prior noise ex4, include = FALSE}
loglike.flat.prior.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

## Model parameters and data

@@

```
## Model predictions

### Pragmatic model

@@

### Prior only model

@@

### Flat prior model

@@

## Compare model predictions to data

@@

## Model comparison

@@

#### Pragmatic model

@@

var steps = 5000
var samples = 1
var mll = AIS(model, {steps:steps, samples: samples})
mll
'
```

@@

```{r bayes factors ex3 children, include = T}
child_ex3_bf <- child_ex3_model_comparison%>%
  group_by(parameter)%>%
  spread(model, logP) %>%
  mutate("pragmatic > flat_prior" = pragmatic - flat_prior,
         "pragmatic > prior_only" = pragmatic - prior_only,
         "flat_prior > prior_only" = flat_prior - prior_only) %>%
  select(-pragmatic,-flat_prior, -prior_only)

>>>>

---
title: "Predicting information integration in pragmatic word learning across development"
subtitle: "Probabilistic Models"
author: "Manuel Bohn, Michael Henry Tessler, Megan Merrick and Michael C. Frank"
output: 
  bookdown::html_document2:
    toc: yes
    toc_float: true
    fig_caption: yes
    code_folding: show
    number_sections: false
  bookdown::pdf_document2:
    toc: yes 
    number_sections: false  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, include = T)

@@

```
# Overview

This file includes model code for the probabilistic models. The section [Model code](#model-code) gives the code for the basic integration model. Firs, we define the objects and locations that feature in the experiments along with some helper functions and then we specify the actual cognitive model.

The section [Adults](#adults) presents the analysis code for Experiment 3 and 4. In the section [Inferring speaker optimality parameter and setting priors](#parameters), we estimated the speaker optimality parameter $\alpha$ based on the data from Experiment 1 and define the prior distribution over objects, $P(r)$, based on Experiment 2.

Then follows the code with which we generated the model predictions for the models in Experiment 3, that is, the integration model (with and without a noise parameter), the no informativeness model and the no common ground model (Section: [Model predictions](#ex3-predictions)). Next, we evaluate the noise parameter for each model and plot the model predictions and correlate predictions and the data for each model (Section: [Compare model predictions to data](#ex3-correlations)). Finally, we compute the log-likelihoods for each model and the Bayes Factors for the model comparisons (Section: [Model comparison](#ex3-comparison)).

We start the section on [Experiment 4](#experiment-4) with specifying $P(r)$ for each version of prior strength manipulations. Next, we generate predictions for the three models (Section: [Model predictions (including noise parameter)](#ex4-predictions), inspect the corresponding noise parameters, and correlate the model predictions with the data (Section: [Compare model predictions to data](#ex4-correlations)). Finally, we compute the log-likelihoods and Bayes Factors (Section: [Model comparison](#ex4-comparison)).

The section [Children](#children) follows the same logic as the one for adults. We first present the code used to estimate the intercept and slope necessary to compute $\alpha$ (using the data from Experiment 5) and $P(r)$ (using the data from Experiment 6) in an age-sensitive way (Section: [Model parameters and data](#child-parameters)). Then we generate model predictions for the three models, each in three versions,: without a noise parameter, with a single noise parameter and with a developmentally changing noise parameter (Section: [Model predictions](#child-predictions)). Then we compare the different noise parameters and correlate model predictions with the data (Section: [Compare model predictions to data](#child-correlations)). Finally, we compute the log-likelihoods for the different models and compare them via Bayes Factors (Section: [Model comparison](#child-comparison)).

@@

#### Inferring speaker optimality parameter and setting priors {#parameters}

@@

### Model predictions {#ex3-predictions}

#### Integration model

@@

```{r predictions pragmatic model ex 3 noise, include = T}
# ex3_pragm_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.rsa.ex3.noise, sep='\n'),
#   data = list(allData = adult_ex3_data, allParams = model_params_ex3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex3_pragm_model_noise.rds")

@@

#### No informativeness model

@@

```{r model prior only noise ex3, include = T}
predictions.prior.only.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

```{r predictions prior only model ex 3 noise, include = T}
# ex3_prior_only_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.prior.only.ex3.noise, sep='\n'),
#   data = list(allData = adult_ex3_data, allParams = model_params_ex3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex3_prior_only_model_noise.rds")

@@

#### No common ground model

@@

```{r predictions flat prior model ex 3 noise, include = T}
# ex3_flat_prior_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.flat.prior.ex3.noise, sep='\n'),
#   data = list(allData = adult_ex3_data, allParams = model_params_ex3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex3_flat_prior_model_noise.rds")

@@

#### Compare model predictions to data {#ex3-correlations}

@@

### Model comparison {#ex3-comparison}

@@

##### Integration model

@@

```{r loglike rsa noise ex3, include = T}
loglike.rsa.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

ex3_pragm_model_noise_loglike <- readRDS("../../stats/saves/ex3_pragm_model_noise_loglike.rds")%>%
  mutate(model = "pragmatic_noise")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))
```
##### No informativeness model

@@

```{r loglike prior only noise ex3, include = T}
loglike.prior.only.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

##### No common ground model

@@

```{r loglike flat prior noise ex3, include = T}
loglike.flat.prior.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

### Model predictions (including noise parameter) {#ex4-predictions}

#### Integration model

```{r model rsa noise ex4, include = T}
predictions.rsa.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

```{r predictions pragmatic model ex 4 noise, include = T}
# ex4_pragm_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.rsa.ex4.noise, sep='\n'),
#   data = list(allData = adult_ex4_data, allParams = model_params_ex4),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex4_pragm_model_noise.rds")

@@

#### No informativeness model

```{r model prior only noise ex4, include = T}
predictions.prior.only.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

```{r predictions prior only model ex 4 noise, include = T}
# ex4_prior_only_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.prior.only.ex4.noise, sep='\n'),
#   data = list(allData = adult_ex4_data, allParams = model_params_ex4),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex4_prior_only_model_noise.rds")

@@

#### No common ground model

```{r model flat prior noise ex4, include = T}
predictions.flat.prior.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

```{r predictions flat prior model ex 4 noise, include = T}
# ex4_flat_prior_model_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.flat.prior.ex4.noise, sep='\n'),
#   data = list(allData = adult_ex4_data, allParams = model_params_ex4),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex4_flat_prior_model_noise.rds")

@@

### Compare model predictions to data {#ex4-correlations}

@@

### Model comparison {#ex4-comparison}

@@

##### Integration model 
```{r loglike rsa noise ex4, include = T}
loglike.rsa.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

##### No informativeness model

```{r loglike prior only noise ex4, include = T}
loglike.prior.only.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

##### No common ground model

```{r loglike flat prior noise ex4, include = T}
loglike.flat.prior.ex4.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

## Model parameters and data {#child-parameters}

@@

```
## Model predictions {#child-predictions}

### Integration model

@@

### No informativeness model

@@

### No common ground model

@@

## Compare model predictions to data {#child-correlations}

@@

## Model comparison {#child-comparison}

@@

#### Integration model

@@

var steps = 5000
var samples = 100
var mll = AIS(model, {steps:steps, samples: samples})
mll
'
```

@@

```{r bayes factors ex3 children, include = T}
child_ex3_bf <- child_ex3_model_comparison%>%
  group_by(parameter)%>%
  spread(model, logP) %>%
  mutate("pragmatic > flat_prior" = exp(pragmatic - flat_prior),
         "pragmatic > prior_only" = exp(pragmatic - prior_only),
         "flat_prior > prior_only" = exp(flat_prior - prior_only)) %>%
  select(-pragmatic,-flat_prior, -prior_only)
